---
output: 
  html_document:
    theme: journal
params: 
  character: 
    value: "Luke Skywalker"
    label: "Select Character"
    choices: !r purrr::map_chr(readRDS("people.Rds"), "name")
    input: select
# character: "Luke Skywalker"  
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(purrr)
library(ggplot2)
library(stringr)
library(rwars)
library(ggrepel)
library(readr)
library(plotly)
library(tibble)
source("flickr_api.R")
people <- readRDS("people.Rds")
```



```{r getpeopledata, eval = FALSE}
# this is how we get the option for the parameters
getDataRobust <- function(ids, f) {
   
   getSingle <- function(id, f) { 
     tryCatch(f(id), error = function(e) { 
      NULL
     })
   }
   
   all <- map(ids, ~getSingle(.x, f))
   not_null <- all[!sapply(all, is.null)]
   
   not_null
}


ids <- 1:100
people <- getDataRobust(ids, get_person)
saveRDS(people, file = "people.Rds")
```


# `r params$character` Quick Facts

```{r flickr_photo}
img_url <- flickr_photo_url(flickr_photos_search_one(paste0(params$character, " lego starwars")))
```

![](`r img_url`)

```{r getcharacterdata}
names <- people %>% 
  map_chr("name")
id <- which(names == params$character)
char <- people[[id]]
```

## Appearance

```{r appearance_species}
species_id <- str_extract(char$species[[1]], "\\d")
species <- get_species(species_id)
species_home_id <-   str_extract(species$homeworld[[1]], "\\d")
species_home <- get_planet(species_home_id)
```

`r params$character` is a `r char$gender` `r species$name`. Their average lifespan is `r species$average_lifespan`, they are `r species$designation` `r species$classification`s originally from `r species_home$name`. 


`r species$name` | `r params$character`
----|-----
Hair | `r char$hair_color`
Eyes | `r char$eye_color`
Mass | `r char$mass`
Height | `r char$height`
Skin |  `r char$skin_color`

## Homeworld

```{r homeworld}
home_id <-   str_extract(char$homeworld[[1]], "\\d")
home <- get_planet(home_id)
```

`r params$character` is from `r home$name`.

```{r flickr_photo2}
img_url <- flickr_photo_url(flickr_photos_search_one(paste0("starwars ", home$name)))
```

![](`r img_url`)

Home Planet | `r home$name`
----|-----
Length of Day | `r home$rotation_period`
Length of Year | `r home$orbital_period`
Climate | `r home$climate`
Gravity | `r home$gravity`
Terrain |  `r home$terrain`

```{r homeworld2}
residents <- map_chr(home$residents, 
                     ~ str_extract(.x, "\\d") %>%
                         get_person() %>%
                         .$name) %>% 
                      unique()
```

Other characters from `r home$name` include: `r residents[which(residents != params$character)]`

## Films

```{r getfilms}
films <- map_chr(char$films, 
                     ~ str_extract(.x, "\\d") %>%
                         get_film() %>%
                         .$title) %>% 
                      unique()
```

`r params$character` appears in: 

`r films`


